A thesis draft

### 2018-10-18 res/sections/07-Introduction.tex
Add vibes logo, proof reading, add content

 Through the history of internet, many factors changed and new innovations came,
 but at the end of the day a foundamental key remained: connection performances
 are, indeed, what the users really want.\todo{Idea: talk about how TCP/IP
   performes quite well in wired connections, while in wireless one, such as
   wireless/satellite, sucks ass}
With the arrival of mobile connectivity, wireless networks started to have an 
important role in everyone's life. In this context network efficiency plays 
an important role with the development of new connectivity standards (e.g 2G, 
3G, LTE) and the creation of common, ad-hoc versions, transmission protocols 
ad-hoc versions (i.e. TCP versions specialized for exploit the characteristic of 
wireless networks). It is for this reason that, beside improved radio 
communications, backend infrastructure began to be refined too. During their 
communication to the destination packets are elaborated, compressed and 
processed in order to make their transmission as fast as possible.
On top of that, mobile traffic is continuously surging every year, with the 
backbones having more and more data to deliver, making in-hardware backbone 
solutions less efficent. Their time-to-deploy, in fact, is very high compared 
to the requirements of the network. With this trend, in-hardware solutions 
will not be able to keep up. Is with this considerations that software a 
software approach is being proposed to become the new way to deploy network 
components. While less efficent, they are faster to bring them up, offering an 
unprecedented flexibility\todo{CITE: SDN/NFV-Based Mobile Packet Core Network 
Architectures: A Survey}.
 Furthermore, new virtualization technologies such as Cloud computing and Docker
 are becoming more and more mature for being employed in large-scale production
 environments, where new frameworks, called orchestrators (e.g. Kubernetes,
 Docker Swarm), allow to easily manage the application lifecyle, i.e.
 deployment, scale up/down and finally, removal.

@@

 This thesis is organized as follows:
 \begin{itemize}
  \item Chapter~\ref{chap:background} describes background information about 5G 
and the technologies used to build an ETSI compliant architecture
  \item Chapter~\ref{chap:prjan} explains the analysis performed on the 
requirements provided and the design choices performed.
  \item Chapter~\ref{chap:archimpl} provides an in-depth view of the 
high-level architectural implementation.
  \item Chapter~\ref{chap:conclusions} wraps up the overall project experience.
 \end{itemize}

>>>>

 Through the history of internet, many factors changed and new innovations came,
 but at the end of the day a foundamental key remained: connection performances
 are, indeed, what the users really want.\todo{Idea: talk about how TCP/IP
   performes quite well in wired connections, while in wireless one, such as
   wireless/satellite, sucks ass}
With the arrival of mobile connectivity, wireless networks started to have an 
important role in everyone's life. In this context network efficiency plays 
an important role with the development of new connectivity standards (e.g 2G, 
3G, LTE) and the creation of common, ad-hoc versions, transmission protocols 
ad-hoc versions (i.e. TCP versions specialized for exploit the characteristic of 
wireless networks). It is for this reason that, beside improved radio 
communications, backend infrastructure began to be refined too. During their 
communication to the destination packets are elaborated, compressed and 
processed in order to make their transmission as fast as possible.
On top of that, mobile traffic is continuously surging every year, with the 
backbones having more and more data to deliver, making in-hardware solutions 
less efficent. Their time-to-deploy, in fact, is very high compared 
to the requirements of the network. With this trend, in-hardware solutions 
will not be able to keep up the Internet needs. Is with this considerations 
that software-based approach, based on virtualization technologies, is being 
proposed to become the new way to deploy network components. While less 
efficent, they are faster to bring them up, offering an unprecedented 
flexibility\todo{CITE: SDN/NFV-Based Mobile Packet Core Network Architectures: A 
Survey}.
 Furthermore, new virtualization technologies such as Cloud computing and Docker
 are becoming more and more mature for being employed in large-scale production
 environments, where new frameworks, called orchestrators (e.g. Kubernetes,
 Docker Swarm), allow to easily manage the application lifecyle, i.e. its
 deployment, scale up/down and finally, removal.

@@

 This thesis is organized as follows:
 \begin{itemize}
  \item Chapter~\ref{chap:background} describes background information about 5G 
and the technologies used to build an ETSI compliant architecture
  \item Chapter~\ref{chap:prjan} explains the analysis performed on the 
requirements provided and the design choices performed.
  \item Chapter~\ref{chap:archimpl} provides an in-depth view of the 
high-level architectural implementation.
  \item Chapter~\ref{chap:conclusions} wraps up the overall project experience 
and takes into account final considerations.
 \end{itemize}

### 2018-10-18 res/sections/08-Background.tex
Change column width in table
Add vibes logo, proof reading, add content

In this chapter we are going to explain the most important concepts of the 
upcoming 5G, talking about the Virtual Network Functions (VNFs) and how first 
experiments with this new technology already begun. Finally we are going to 
briefly introduce and explain the role of the technologies used in this 
Thesis.

\section{The upcoming connectivity standard: 5G}
The continuous innovation in the mobile network connectivity is leading to the
creation of a new standard, the 5G, which is estimated to arrive to the market
constumer in 2020\todo{CITE}. Lead by the Next Generation Mobile Network (NGMN)
alliace, a group composed by the major players in the field of mobile
connectivity, the 5G aims to offer at the end-user a new way to browse the web,
download and watch interactive content. Relatively to 4G, 5G points to have data
rates $10$ times better, with $10$ times smaller end-to-end latency and an
increased connection densisty by $100$ times\todo{CITE}.

\begin{table}[t]
\centering
\resizebox{\textwidth}{!}{%
  \begin{tabular}{p{4cm}|p{5,5cm}|p{5,5cm}}
\textbf{Attribute}                                                    & 
\textbf{LTE capability}                                                          
                                                                           & 
\textbf{Improvement needed to meet NGMN requirements}                 \\ \hline
\textbf{Data rate (per user)}                                         & Up to 
100 Mb/s on average Peaks of 600 Mb/s (Cat 11/12).                               
                                                                     & 10X 
expected on average and peak rates and 100X expected on cell edge \\
\textbf{End-to-end latency}                                           & 10 ms 
for two-way RAN (pre-scheduled). Typically, up to 50 ms end-to-end if other 
factors are considered (e.g., transmission, CN, internet, proxy servers). & 10X 
(smaller)                                                         \\
\textbf{Mobility}                                                     & 
Functional up to 350 km/h (for certain bands up to 500 km/h). No support for 
civil aviation.                                                                & 
1.5X                                                                  \\
\textbf{\begin{tabular}[c]{@{}l@{}}Connection\\ density\end{tabular}} & 
Typically $\sim$2,000 active users/km2.                                          
                                                                           & 
100X                                                                 
\end{tabular}%
}
\caption[5G improvement over LTE]{An extract from the official 5G white paper 
illustating the improvements of 5G relatively to LTE connections.}
\label{chap:intro:table:ltevs5g}
\end{table}

Part of the new requirements can be satisfied using a large radio spectrum with
higher frequencies. The utilization of higher frequencies, though, mean that the
radio signals can be easly disrupted by physical objects, like buildings and
many geographical elements (such as hills and mountains), clashing with the
expectation of an ever-reachable connectivity. It is here where virtualization
plays an important role. In fact, the re-design of some network components today
existing via hardware can transform a monolitic networking approach to a modular
one, exploting the flexibility that Virtual Network Functions (VNFs)\todo{CITE
  VNF-P: A Model for Efficient Placement of Virtualized Network Functions} can
offer thanks to virtualization, closing the gap to the use-case fullfilment
defined by the NGMS alliance\todo{CITE: NGMN View on 5G architecture}.

@@

\subsubsection{Network slicing}
The role of a ``network slice'' in a 5G architecture is to specificatly handle a
Control-place of a praticular service (e.g. smartphones traffic, autonomous
driving, massive IoT), deploying resources in a manner that assure the required
latency, security and reliability. While some legacy of very peculiar services
could require specific hardware, the common resources between services could be
shared in a virtualized way, providing auto-scaling capabilities in services
that are under heavy network pressure.

@@

>>>>

In this chapter we are going to explain the most important concepts of the 
upcoming 5G standard, talking about the Virtual Network Functions (VNFs) and 
how first experiments with this new technology already begun. Finally we are 
going to briefly introduce and explain the technologies involved and their role.

\section{The upcoming connectivity standard: 5G}
The continuous innovation in the mobile network connectivity is leading to the
creation of a new standard, the 5G, which is estimated to arrive to the market
constumer in 2020\todo{CITE}. Lead by the Next Generation Mobile Network (NGMN)
alliace, a group composed by the major players in the field of mobile
connectivity, the 5G aims to offer not only at the end-user a new way to browse 
the web,
download and watch interactive content but also to create an ad-hoc solution 
for Machine-to-Machine (M2M) data traffic, which is increasing more than ever 
thanks to the spreading of IoT devices, which sensors need to continuosly send 
data to servers/datacenters. Relatively to LTE\todo{Explain acronym}, 5G points 
to have data rates $10$ times better, with $10$ times smaller end-to-end 
latency and an increased connection densisty by $100$ times\todo{CITE}.

\begin{table}[t]
\centering
\resizebox{\textwidth}{!}{%
  \begin{tabular}{p{4,5cm}|p{5,5cm}|p{5cm}}
\textbf{Attribute}                                                    & 
\textbf{LTE capability}                                                          
                                                                           & 
\textbf{Improvement needed to meet NGMN requirements}                 \\ \hline
\textbf{Data rate (per user)}                                         & Up to 
100 Mb/s on average Peaks of 600 Mb/s (Cat 11/12).                               
                                                                     & 10X 
expected on average and peak rates and 100X expected on cell edge \\
\textbf{End-to-end latency}                                           & 10 ms 
for two-way RAN (pre-scheduled). Typically, up to 50 ms end-to-end if other 
factors are considered (e.g., transmission, CN, internet, proxy servers). & 10X 
(smaller)                                                         \\
\textbf{Mobility}                                                     & 
Functional up to 350 km/h (for certain bands up to 500 km/h). No support for 
civil aviation.                                                                & 
1.5X                                                                  \\
\textbf{\begin{tabular}[c]{@{}l@{}}Connection\\ density\end{tabular}} & 
Typically $\sim$2,000 active users/km2.                                          
                                                                           & 
100X                                                                 
\end{tabular}%
}
\caption[5G improvement over LTE]{An extract from the official 5G white paper 
illustating the improvements of 5G relatively to LTE connections.}
\label{chap:intro:table:ltevs5g}
\end{table}

Part of the new requirements can be satisfied using a large radio spectrum with
higher frequencies. The utilization of higher frequencies, though, mean that the
radio signals can be easly disrupted by physical objects, like buildings and
many geographical elements (such as hills and mountains), clashing with the
expectation of an ever-reachable connectivity. It is here where virtualization
plays an important role. In fact, the re-design of some network components today
existing via hardware can transform a monolitic networking approach to a modular
one, exploting the flexibility that Virtual Network Functions (VNFs)\todo{CITE
  VNF-P: A Model for Efficient Placement of Virtualized Network Functions} can
offer thanks to virtualization, closing the gap to the use-case fullfilment
defined by the NGMS alliance\todo{CITE: NGMN View on 5G architecture} which 
require a descreased time to set up and deploy network services (specificatly, 
from 90 hours to 90 minutes\todo{Search this requirement, check 90h 'cause 
I'm not sure about it}).

@@

\subsubsection{Network slicing}
The role of a ``network slice'' in a 5G architecture is to specificatly handle a
Control-place of a praticular service (e.g. smartphones traffic, autonomous
driving, massive IoT), deploying resources in a manner that assure the required
latency, security and reliability. While some very peculiar legacy services
could require specific hardware, the common resources between services could be
shared in a virtualized way, providing auto-scaling capabilities in services
that are under heavy network pressure.

@@

\begin{figure}[t]
 \centering
 \includegraphics[scale=1]{vibes_logo}
 \caption{ViBeS project logo}
 \label{chap:background:img:vibes_logo}
\end{figure}
